### 2017-11-04 submission/manuscript.Rmd
incorporate n into manuscript and incorporate tables into methods

**Background.** An increasing body of literature suggests that there a role for the microbiota in colorectal cancer. Important players within this axis have ranged from individual microbes to the whole community. A recent meta-analysis investigated this but only focused on potential biomarkers. This study expands on this previous research and tests the hypothesis that detectable changes in the bacterial community are important both to increasing relative risk and model accuracy for adenoma and carcinoma. To test this hypothesis we examined both fecal and tissue samples from 2109 individuals across 14 different studies.

@@

Our analysis found a continuous decreae in Shannon diversity from control to adenoma to carcinoma and a significantly increased relative risk for carcinoma with lower diversity. Using the CRC associated genera only this relative risk was higher than Shannon diversity. However, adding the community context in which these CRC associated genera are present increases prediction models AUC. Although we analyze a data set with a large number of total individuals each individual study was underpowered for effect size differences of 10% or below between the case and control.

@@

***Data Set Breakdown:***
In total there were 7 studies with only fecal samples (Ahn, Baxter, Brim, Hale, Wang, Weir, and Zeller), 5 studies with only tissue samples (Burns, Dejea, Geng, Lu, Sanapareddy), and 2 studies with both fecal and tissue samples (Chen and Flemer). The total number of individuals initially run through the sequence processing for the fecal samples was 1899 and for the tissue samples was 462.

>>>>

**Background.** An increasing body of literature suggests that there a role for the microbiota in colorectal cancer. Important players within this axis have ranged from individual microbes to the whole community. A recent meta-analysis investigated this but only focused on potential biomarkers. This study expands on this previous research and tests the hypothesis that detectable changes in the bacterial community are important both to increasing relative risk and model accuracy for adenoma and carcinoma. To test this hypothesis we examined both feces (total individuals = `r sum(stool_n$control) + sum(stool_n$adenoma, na.rm = T) + sum(stool_n$cancer, na.rm = T)`) and tissue (total samples = `r sum(tissue_n$control) + sum(tissue_n$adenoma, na.rm = T) + sum(tissue_n$cancer, na.rm = T)`) across 14 different studies.

@@

Our analysis found a continuous decrease in Shannon diversity from control to adenoma to carcinoma and a significantly increased relative risk for carcinoma with lower diversity. Using the CRC associated genera only this relative risk was higher than Shannon diversity. However, adding the community context in which these CRC associated genera are present increases prediction models AUC. Although we analyze a data set with a large number of total individuals each individual study was underpowered for effect size differences of 10% or below between the case and control.

@@

***Data Set Breakdown:***
In total there were 7 studies with only fecal samples (Ahn, Baxter, Brim, Hale, Wang, Weir, and Zeller), 5 studies with only tissue samples (Burns, Dejea, Geng, Lu, Sanapareddy), and 2 studies with both fecal and tissue samples (Chen and Flemer). The total number of individuals that were analyzed after sequence processing for feces was `r sum(stool_n$control) + sum(stool_n$adenoma, na.rm = T) + sum(stool_n$cancer, na.rm = T)` [Table 1]. The total number of matched and unmatched tissue samples that were analyzed after sequence processing was `r sum(tissue_n$control) + sum(tissue_n$adenoma, na.rm = T) + sum(tissue_n$cancer, na.rm = T)` [Table 2].

### 2017-11-04 submission/manuscript.tex
incorporate n into manuscript and incorporate tables into methods

\textbf{Background.} An increasing body of literature suggests that
there a role for the microbiota in colorectal cancer. Important players
within this axis have ranged from individual microbes to the whole
community. A recent meta-analysis investigated this but only focused on
potential biomarkers. This study expands on this previous research and
tests the hypothesis that detectable changes in the bacterial community
are important both to increasing relative risk and model accuracy for
adenoma and carcinoma. To test this hypothesis we examined both fecal
and tissue samples from 2109 individuals across 14 different studies.

@@

Our analysis found a continuous decreae in Shannon diversity from
control to adenoma to carcinoma and a significantly increased relative
risk for carcinoma with lower diversity. Using the CRC associated genera
only this relative risk was higher than Shannon diversity. However,
adding the community context in which these CRC associated genera are
present increases prediction models AUC. Although we analyze a data set
with a large number of total individuals each individual study was
underpowered for effect size differences of 10\% or below between the
case and control.

@@

\textbf{\emph{Data Set Breakdown:}} In total there were 7 studies with
only fecal samples (Ahn, Baxter, Brim, Hale, Wang, Weir, and Zeller), 5
studies with only tissue samples (Burns, Dejea, Geng, Lu, Sanapareddy),
and 2 studies with both fecal and tissue samples (Chen and Flemer). The
total number of individuals initially run through the sequence
processing for the fecal samples was 1899 and for the tissue samples was
462.

>>>>

\textbf{Background.} An increasing body of literature suggests that
there a role for the microbiota in colorectal cancer. Important players
within this axis have ranged from individual microbes to the whole
community. A recent meta-analysis investigated this but only focused on
potential biomarkers. This study expands on this previous research and
tests the hypothesis that detectable changes in the bacterial community
are important both to increasing relative risk and model accuracy for
adenoma and carcinoma. To test this hypothesis we examined both feces
(total individuals = 1737) and tissue (total samples = 492) across 14
different studies.

@@

Our analysis found a continuous decrease in Shannon diversity from
control to adenoma to carcinoma and a significantly increased relative
risk for carcinoma with lower diversity. Using the CRC associated genera
only this relative risk was higher than Shannon diversity. However,
adding the community context in which these CRC associated genera are
present increases prediction models AUC. Although we analyze a data set
with a large number of total individuals each individual study was
underpowered for effect size differences of 10\% or below between the
case and control.

@@

\textbf{\emph{Data Set Breakdown:}} In total there were 7 studies with
only fecal samples (Ahn, Baxter, Brim, Hale, Wang, Weir, and Zeller), 5
studies with only tissue samples (Burns, Dejea, Geng, Lu, Sanapareddy),
and 2 studies with both fecal and tissue samples (Chen and Flemer). The
total number of individuals that were analyzed after sequence processing
for feces was 1737 {[}Table 1{]}. The total number of matched and
unmatched tissue samples that were analyzed after sequence processing
was 492 {[}Table 2{]}.
